---
title: "CHEK2_UKB_LM"
output: html_document
date: "2025-03-06"
---

```{r}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  tidyverse,    # data management + ggplot2 graphics, 
  stringr,      # manipulate text strings 
  purrr,        # loop over objects in a tidy way
  gtsummary,    # summary statistics and tests 
  broom,        # tidy up results from regressions
  lmtest,       # likelihood-ratio tests
  parameters,   # alternative to tidy up results from regressions
  see,          # alternative to visualise forest plots
  readxl
)
```


```{r}
pheno_ukb<-read_excel("R:/Medicine/Hematology-Oncology/Godley_Lab/UK_Biobank/CHEK2/Proteomics/Phenotype_UKB_Minus_Cancer_risk.xlsx")

model_df<- pheno_ukb[, c("eid", "Sex", "Age_at_recruitment","Malignant_Cancer_Status","Age_at_First_Cancer","Cancer_Type","Heme_Type","First_Cancer_Code")]


# Rename columns by index
colnames(model_df)[colnames(model_df) == "Cancer_Type_x"] <- "Cancer_Type"
```

#Read_CHEK2_Status
```{r}
chek2_pids<-read.csv("R:/Medicine/Hematology-Oncology/Godley_Lab/UK_Biobank/CHEK2/Proteomics/CHEK2_alleles_PLP.csv")
# Add a new column 'CHEK2_Status' where the value is 'Yes' if EID is in the chek2_eids list
model_df$CHEK2_Status <- ifelse(model_df$eid %in% unique(chek2_pids$Participant.ID), 1, 0)
table(model_df$CHEK2_Status)
```



```{r}
library(dplyr)



# Create binary columns for each Cancer_Type category
model_df <- model_df %>%
  mutate(
    Heme_malignancy_Status = ifelse(Cancer_Type == 'Heme_Malignancy', 1, 0),
    Solid_tumor_Status = ifelse(Cancer_Type == 'Solid_Tumor', 1, 0),
    Other_Malignancy = ifelse(Cancer_Type == 'Other_Malignancy_Secondary', 1, 0)
  )

# Create binary columns for each Cancer_Type category
model_df <- model_df %>%
  mutate(
    Myeloid_Status = ifelse(Heme_Type == 'Myeloid', 1, 0),
    Lymphoid_Status = ifelse(Heme_Type == 'Lymphoid', 1, 0),
      )


```


Category

```{r}

icd10_dict<-read_excel("R:/Medicine/Hematology-Oncology/Godley_Lab/UK_Biobank/CHEK2/Proteomics/Coding/Merged_ICD10_Categories.xlsx")

model_df$First_Cancer_Code_Broad <- ifelse(is.na(model_df$First_Cancer_Code), NA, substr(model_df$First_Cancer_Code, 1, 3))
icd_broad_unique<-icd10_dict[, c("ICD10_Broad", "Description_Broad")]
icd_broad_unique <- unique(icd_broad_unique)

# Merge with icd10_dict to get Description_Broad
model_df <- merge(model_df, icd_broad_unique[, c("ICD10_Broad", "Description_Broad")], 
            by.x = "First_Cancer_Code_Broad", by.y = "ICD10_Broad",all.x = TRUE)

model_df$Description_Broad[model_df$Description_Broad == "Carcinoma in situ of breast"] <- "Malignant neoplasm of breast"


```



#Smoking Status
```{r}
smoking_status<-read.csv("R:/Medicine/Hematology-Oncology/Godley_Lab/UK_Biobank/CHEK2/Proteomics/Smoking_Status_UKB.csv")
smoking_status <- smoking_status %>%
  mutate(Smoking_Status = ifelse(rowSums(select(., starts_with("p20160")) == 1, na.rm = TRUE) > 0, 1, 0))

table(smoking_status$p20160_i0)
model_df<-merge(model_df,smoking_status,by="eid",how="inner")

# Drop specific columns by name
model_df <- model_df[, !names(model_df) %in% c("p21022","p31","p20160_i0","p20160_i1","p20160_i2", "p20160_i3")]



```

#Age of Onset of Cancer

```{r}
# Rename columns by index
colnames(model_df)[colnames(model_df) == "p34"] <- "Birth_Year"

# Assuming 'birth_year' is the column containing birth years
model_df$current_age <- 2025 - model_df$Birth_Year


# Assuming the columns are 'Malignant_Status', 'Age_at_recruitment', and 'Age_at_first_cancer'
model_df$Age <- ifelse(model_df$Malignant_Cancer_Status == FALSE, model_df$Age_at_recruitment, model_df$Age_at_First_Cancer)
```


#Convert Malignant Cancer to Binary
```{r}
# Convert "TRUE"/"FALSE" to 1/0
model_df$Malignant_Cancer_Status <- ifelse(model_df$Malignant_Cancer_Status == "TRUE", 1, ifelse(model_df$Malignant_Cancer_Status == "FALSE", 0, NA))
```


#Convert Sex to Binary
```{r}
# Convert "Male"/"Female" to 1/0
model_df$Sex <- ifelse(model_df$Sex == "Male", 1, ifelse(model_df$Sex == "Female", 0, NA))
table(model_df$Sex)
```


#Create age decile 0-10,10-20,20-30 etc
```{r}
model_df <- model_df %>%
  mutate(age_cat = cut(Age, 
                       breaks = seq(0, 100, by = 10), 
                       labels = paste0(seq(10, 100, by = 10), "s"), 
                       include.lowest = TRUE, 
                       right = FALSE))
```


```{r}


# Convert to binary form
model_df_wide <- model_df %>%
  mutate(Present = 1) %>%  # Add a column with 1 for presence
  pivot_wider(names_from = Description_Broad, values_from = Present, values_fill = list(Present = 0))


model_df_wide$Lymphoid_Status[is.na(model_df_wide$Lymphoid_Status)] <- 0
model_df_wide$Myeloid_Status[is.na(model_df_wide$Myeloid_Status)] <- 0
model_df_wide$Heme_malignancy_Status[is.na(model_df_wide$Heme_malignancy_Status)] <- 0
model_df_wide$Solid_tumor_Status[is.na(model_df_wide$Solid_tumor_Status)] <- 0


```





#Explanatory Variables/ Co-variates
```{r}
explanatory_vars<-c("Sex","Age","CHEK2_Status","Smoking_Status")


## drop rows with missing information for variables of interest 
model_df_final <- model_df %>% 
  drop_na(any_of(explanatory_vars))


# Check for NA values in the explanatory variables columns
na_positions <- model_df %>%
  select(any_of(explanatory_vars)) %>%
  is.na()

table(model_df_final$CHEK2_Status)




```






#Test glm with one Variable CHEK2 Status


```{r}
model_df_final %>% 
  mutate(CHEK2_Status = as.factor(CHEK2_Status)) %>%  # Convert to factor
  glm(formula = Malignant_Cancer_Status ~ CHEK2_Status, family = binomial, data = .) %>% 
  summary()
```
#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_Cancer_Status ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_final)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
```


```{r}
#

univ_tab <- model_df_final %>% 
  dplyr::select(explanatory_vars, Malignant_Cancer_Status) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_Cancer_Status,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )
```


```{r}
univ_tab
```


##Multivariate Regression
```{r}

mv_reg <- glm(Malignant_Cancer_Status ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_final)

summary(mv_reg)

```



```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```



```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```


#By Cancer Types
```{r}
explanatory_vars<-c("Sex","Age","CHEK2_Status","Smoking_Status")


## drop rows with missing information for variables of interest 
model_df_wide <- model_df %>% 
  drop_na(any_of(explanatory_vars))


# Check for NA values in the explanatory variables columns
na_positions <- model_df %>%
  select(any_of(explanatory_vars)) %>%
  is.na()

table(model_df_wide$CHEK2_Status)
```


#Check why only 1062 in CHEK2

```{r}

problematic_rows <- model_df %>%
filter(CHEK2_Status == 1 &  Malignant_Cancer_Status == 1 )


blank_rows<-problematic_rows[which(is.na(problematic_rows$Cancer_Type)),]

blank_cancer<-pheno_ukb[pheno_ukb$eid%in%blank_rows$eid,]
```





#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Heme_malignancy_Status ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))


```


```{r}
#

univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Heme_malignancy_Status) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Heme_malignancy_Status,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )
```


```{r}
univ_tab
```



##Multivariate Regression
```{r}

mv_reg <- glm(Heme_malignancy_Status ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

```

```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```



```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```


#Solid Tumor


#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Solid_tumor_Status ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
```


```{r}
#

univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Solid_tumor_Status) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Solid_tumor_Status,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )
```


```{r}
univ_tab
```




##Multivariate Regression
```{r}

mv_reg <- glm(Solid_tumor_Status ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

```


```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```


```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```





#Myeloid Malignancy


#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Myeloid_Status ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
```


```{r}
#

univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Myeloid_Status) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Myeloid_Status,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )
```


```{r}
univ_tab
```

##Multivariate Regression
```{r}

mv_reg <- glm(Myeloid_Status ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

```

```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```



```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```



#Lymphoid Malignancy

#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Lymphoid_Status ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
```


```{r}
#

univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Lymphoid_Status) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Lymphoid_Status,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab
```




##Multivariate Regression
```{r}

mv_reg <- glm(Lymphoid_Status ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

```
```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```


```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```
#Malignant Neoplasm of Breast
```{r}
explanatory_vars<-c("Sex","Age","CHEK2_Status","Smoking_Status")


## drop rows with missing information for variables of interest 
model_df_wide <- model_df_wide %>% 
  drop_na(any_of(explanatory_vars))


# Check for NA values in the explanatory variables columns
na_positions <- model_df %>%
  select(any_of(explanatory_vars)) %>%
  is.na()

table(model_df_wide$CHEK2_Status)


# Replace spaces with underscores
colnames(model_df_wide) <- gsub(" ", "_", names(model_df_wide))

```


#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_neoplasm_of_breast ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
```


```{r}
#

univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Malignant_neoplasm_of_breast) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_neoplasm_of_breast,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab
```

##Multivariate Regression
```{r}

mv_reg <- glm(Malignant_neoplasm_of_breast ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

```


```{r}
pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()
```


```{r}
## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names
```

#Malignant Neoplasm of Prostate

#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_neoplasm_of_prostate ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))


univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Malignant_neoplasm_of_prostate) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_neoplasm_of_prostate,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab
```




##Multivariate Regression
```{r}

mv_reg <- glm(Malignant_neoplasm_of_prostate ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()

## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names

```


#Malignant Neoplasm of Bronchus and Lung

#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_neoplasm_of_bronchus_and_lung ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))


univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Malignant_neoplasm_of_bronchus_and_lung) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_neoplasm_of_bronchus_and_lung,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab


##Multivariate Regression

mv_reg <- glm(Malignant_neoplasm_of_bronchus_and_lung ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()

## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names

```




#Malignant Neoplasm of Thyroid

#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_neoplasm_of_thyroid_gland ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))


univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Malignant_neoplasm_of_thyroid_gland) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_neoplasm_of_thyroid_gland,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab


##Multivariate Regression

mv_reg <- glm(Malignant_neoplasm_of_thyroid_gland ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()

## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names

```








#Malignant Neoplasm of Thyroid

#Univariate
```{r}
models <- explanatory_vars %>%       # begin with variables of interest
  str_c("Malignant_neoplasm_of_colon ~ ", .) %>%         # combine each variable into formula ("outcome ~ variable of interest")
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = model_df_wide)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))


univ_tab <- model_df_wide %>% 
  dplyr::select(explanatory_vars, Malignant_neoplasm_of_colon) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = Malignant_neoplasm_of_colon,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

univ_tab


##Multivariate Regression

mv_reg <- glm(Malignant_neoplasm_of_colon ~ Sex + Age + CHEK2_Status+Smoking_Status,family = "binomial", data = model_df_wide)

summary(mv_reg)

pacman::p_load(easystats)

## remove the intercept term from your multivariable results
mv_reg %>% 
  model_parameters(exponentiate = TRUE) %>% 
  plot()

## choose a model using forward selection based on AIC
## you can also do "backward" or "both" by adjusting the direction
final_mv_reg <- mv_reg %>%
  step(direction = "forward", trace = FALSE)

mv_tab_base <- final_mv_reg %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%  ## get a tidy dataframe of estimates 
  mutate(across(where(is.numeric), round, digits = 2))     

mv_tab <- tbl_regression(final_mv_reg, exponentiate = TRUE)


## combine with univariate results 
tbl_merge(
  tbls = list(univ_tab, mv_tab),                          # combine
  tab_spanner = c("**Univariate**", "**Multivariable**")) # set header names

```

Cumulative Incidence
```{r}
library(ggplot2)
library(dplyr)

cumulative_case_counts <- model_df_final %>% 
  count(Age) %>%                # count of rows per day (returned in column "n")   
  mutate(                         
    cumulative_cases = cumsum(n)       # new column of the cumulative number of rows at each date
    )

plot_cumulative <- ggplot()+
  geom_line(
    data = cumulative_case_counts,
    aes(x = Age, y = cumulative_cases),
    size = 2,
    color = "blue")

plot_cumulative


# Compute cumulative counts for each CHEK2_Status group
cumulative_case_counts <- model_df_final %>%
  group_by(CHEK2_Status, Age) %>%
  count() %>%
  mutate(cumulative_cases = cumsum(n))

# Plot cumulative incidence with different colors for CHEK2_Status
plot_cumulative <- ggplot(data = cumulative_case_counts, aes(x = Age, y = cumulative_cases, color = CHEK2_Status)) +
  geom_line(size = 2) +
  labs(title = "Cumulative Incidence of Cancer by Age",
       x = "Age of Onset",
       y = "Cumulative Cases",
       color = "CHEK2 Status") + 
  theme_minimal()

# Display the plot
plot_cumulative
```


```{r}
library(ggplot2)
library(dplyr)

# Filter for Malignant_Cancer_Status == 1, round Age, and compute cumulative counts
cumulative_case_counts <- model_df_final %>%
  filter(Malignant_Cancer_Status == 1) %>%  # Keep only rows with cancer status = 1
  mutate(Age = round(Age)) %>%  # Round Age to nearest integer
  group_by(CHEK2_Status, Age) %>%
  count() %>%
  arrange(CHEK2_Status, Age) %>%  # Ensure sorting before cumulative sum
  group_by(CHEK2_Status) %>%  # Group by CHEK2_Status before cumsum
  mutate(cumulative_cases = cumsum(n))

# Compute total individuals in each CHEK2_Status group (denominator)
total_counts <- model_df_final %>%
  group_by(CHEK2_Status) %>%
  summarise(total_individuals = n())

# Merge with cumulative counts and normalize
cumulative_case_counts <- cumulative_case_counts %>%
  left_join(total_counts, by = "CHEK2_Status") %>%
  mutate(normalized_cumulative_cases = cumulative_cases / total_individuals)

# Plot cumulative incidence as a proportion of each group with updated legend labels
plot_cumulative <- ggplot(data = cumulative_case_counts, aes(x = Age, y = normalized_cumulative_cases, color = as.factor(CHEK2_Status))) +
  geom_line(size = 1.5) +  # Ensure line plot
  scale_color_manual(name = "CHEK2 Status", 
                     values = c("0" = "blue", "1" = "red"),  # Assign colors
                     labels = c("0" = "CHEK2(WT)", "1" = "CHEK2(P/LP)")) +  # Update legend labels
  labs(title = "Normalized Cumulative Incidence of Cancer by Age",
       x = "Age of Onset",
       y = "Proportion of Affected Individuals") + 
  theme_minimal()

# Display the plot
plot_cumulative

```

